---
title: "ESM 244 Lab 6"
author: "Sara Orofino"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Goals:

- Read in spatial data  
- Simplify polygons   
- Add and transform projections (st_transform, st_crs)  
- Create a bunch of maps  
- Join spatial data 
- Find intersections of spatial data 
- Interactive plots   
- Create sf from lat/long coordinates    

Load packages: 
```{r}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
library(plotly)


```


###Example 1. Dams in California   

Notes for Code chunk:  
- Use layer = "" and specify the beginning of the file path  (whatever is before the .prj, .shp etc.)  
- Specify dplyr and use the select function to reduce the number of attributes to just the ecoregion name attribute and associated spatail information.   
- Use st_simplify() to make files sizes easier to work with (if the ecoregion is within 100 meters of a line it will collapse together onto one line)  
- Transform coordinate reference system (can check current crs using st_crs(ca_eco))  


```{r}

ca_eco <- read_sf(".", layer = "ca_eco") %>% 
  dplyr::select(US_L3NAME) %>% 
  rename(Region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% 
  st_transform(crs = 4326)

```

```{r}

ca_counties <- read_sf(".", layer = "california_county_shape_file") #No existing crs 

# Set crs for ca_counties to WGS84:

st_crs(ca_counties) = 4326   


```

```{r}

ca_dams <- read_sf(".", layer =  "California_Jurisdictional_Dams") %>% 
  rename(Condition = Condition_) 

# Factor Relevel - used to override the default setting of alphabetical ordering 

ca_dams$Condition <- fct_relevel(ca_dams$Condition, "Fair", "Satisfactory", "Unsatisfactory", "Poor")

```


##MAPS!!! 


```{r}

plot(ca_eco)
plot(ca_counties)

```


Make a map  with ggplot!  

```{r}

# RColor Brewer to specify color palette for the 13 ecoregions 
# Vector for number of regions we need

color_count <- 13

# brewer.pal(number of colors in existing ramp, "rampname")(vector for colors)
my_colors  <- colorRampPalette(brewer.pal(10, "Set2"))(color_count)

#Graph code - color = "NA" removes the outline for counties; change color scheme with scale_fill_manual(); add layers using geom_sf() but specify where data comes from! 
# coord_sf(datum = NA) gets rid of lat/long degrees 

ggplot(ca_eco)  +
  geom_sf(aes(fill= Region), 
          color = "NA",
          show.legend = FALSE) +
  scale_fill_manual(values =  my_colors) +
  geom_sf(data = ca_counties,
          fill = "NA",
          color  = "grey30",  
          size =  0.1) +
  geom_point(data = ca_dams, aes(x = Longitude, y = Latitude),
             size = 1, color = "grey10", alpha = 0.5)  +
  theme_minimal() +
  coord_sf(datum  = NA) 

```

###Example 2. Dams in the Sierra Nevada Ecoregion  


```{r}
# Join information in the ecoregions dataframe for Sierra Nevada with information in the dam dataframe

sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% 
  st_join(ca_dams) 

ggplot(sn) +
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") +
  geom_sf(fill = "lemonchiffon4", color = "NA") +
  geom_point(aes(x = Longitude, y = Latitude), size = 0.5, color = "red4")

```


###Example 3. Ecoregions for Santa Barbara County   


```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")

eco_clip <- st_intersection(ca_eco, sb)

ggplot(eco_clip) +
  geom_sf(data = ca_counties, fill = "grey90", color = "grey80",
          size = 0.2) +
  geom_sf(aes(fill = Region, line = "NA")) +
  scale_fill_manual(values = c("darkolivegreen2", "darkolivegreen", "gold2")) +
  coord_sf(xlim = c(-121, -119), ylim = c(33.5,35.5)) +
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2) +
  geom_text(x = -119.6982, y = 34.35, label = "Santa Barbara") +
  theme_minimal() +
  theme(legend.position = c(0.5,0.15)) + 
  labs(x = "", y = "", title = "Santa Barbara County Ecoregions")
  
```


###Example 4. Intro to interactive plots with tmap 

```{r}

map_sb_eco <- tm_shape(eco_clip) +
  tm_fill("Region", palette = "RdPu", alpha = 0.5) +
  tm_shape(ca_counties) +
  tm_borders()

#sets tmap to interactive viewing: 

tmap_mode("view")
map_sb_eco

# Can use leaflet::providers() to view what basemaps are available 
# Setting the basemap will take away the options for different ones in the interactive viewing 

# Example with different basemap 

tm_basemap("CartoDB.DarkMatter") + 
  tm_shape(eco_clip) +
  tm_borders(col = "white")
```

###Example 5. Fault Lines

```{r}
# Separate takes information from the LTYPE column and separates the varaibles in the that column into different columns, tell it the variables are separted with commas by using sep = ","  


fault_lines <- read_sf(".", layer = "GMC_str_arc") %>% 
  st_transform(crs = 4326) %>% 
  separate(LTYPE, into = c("syn_ant", "certainty", "direction", sep = ","))

ggplot() +
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  theme_dark()


# Fault lines in SB county only:

sb_faults <- fault_lines %>% 
  st_intersection(sb)

ggplot() +
  geom_sf(data = sb) +
  geom_sf(data = sb_faults, aes(color = syn_ant))

tm_basemap("CartoDB.DarkMatter") + 
  tm_shape(sb) +
  tm_borders(col = "grey50", lwd = 2) + 
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange", "purple"), lwd = 2)
```



###Example 6. Faceted Maps   

Frequently faceted by year, but here we'll use types of fault lines  

```{r}

ggplot() +
  geom_sf(data  = ca_counties, fill = "black", color= "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  facet_wrap(~syn_ant)

# can use tm_facets to facet wrap in tmaps
```


###Example 7. Making Spatial Points  

```{r}

ca_sites <- read_csv("cadfw_sensitive_sites.csv")

#Specify which columns the lat/long are in using coords() - notice the geometry column added to the end
# Have to specify long then lat


sites_sf <- st_as_sf(ca_sites, coords = c("LONDD", "LATDD"),
                     crs = 4326)

ggplot() + 
  geom_sf(data = ca_counties, fill = "grey40") +
  geom_sf(data = sites_sf, aes(color = PRI_CODE), size = 0.3)

```


###Example 8. Chloropleth of CA Counties by number of dams in each county 

```{r}
# Be sure that the spatial information for dams is within CA counties - find intersection 

intersection <- st_intersection(ca_dams, ca_counties)

# Want to count how many dams there are in each county:

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()

# Then merge to the ca_counties dataframe: 

ca_total <- ca_counties %>% 
  st_join(dams_per_county) %>% 
  dplyr::select(NAME.x, n) %>% 
  rename(name = NAME.x)

# Reassign NA values to zero:
ca_total$n[is.na(ca_total$n)] <- 0


# Graph it 

ggplot() +
  geom_sf(data =ca_total, aes(fill = n), size = 0.2) +
  theme_minimal() +
  scale_fill_continuous(low = "yellow", high = "red")

```



